---
title: "Formal Languages and Automata"
author: "David J. Hunter"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(metathis)
meta() %>%
  meta_description(
    "Slides, assignments, and code for my class on the theory of computation: CS/MA-135: Formal Languages and Automata."
  ) %>%
  meta_name("github-repo" = "djhunter/cryptography") %>%
  meta_viewport() %>%
  meta_social(
    title = "Cryptography and Coding Theory",
    url = "https://djhunter.github.io/cryptography/index.html",
    image = "https://upload.wikimedia.org/wikipedia/en/5/5b/Codes_And_Keys_Death_Cab_For_Cutie.jpg",
    image_alt = "Codes and Keys album art",
    og_author = c("David J. Hunter")
  )
```

This site provides slides, assignments, and other resources that I use in my class on the theory of computation (CS/MA-135: Formal Languages and Automata, Westmont College).

# Slides 

1. []() 

---
# 22. [Reed Solomon Codes](slides/22codes6.html)
# 23. [Codes and Encryption](slides/23codes7.html)
---

# Prework

1. []()

# Other Resources

- [Syllabus](syllabus/csma135hunterS24.pdf) (PDF)

All of the pages and slides on this site were authored in RMarkdown. The source code is available on GitHub: https://github.com/djhunter/computation

