---
title: "Formal Languages and Automata"
author: "David J. Hunter"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(metathis)
meta() %>%
  meta_description(
    "Slides, assignments, and code for my class on the theory of computation: CS/MA-135: Formal Languages and Automata."
  ) %>%
  meta_name("github-repo" = "djhunter/computation") %>%
  meta_viewport() %>%
  meta_social(
    title = "Formal Languages and Automata",
    url = "https://djhunter.github.io/computation/index.html",
    image = "https://upload.wikimedia.org/wikipedia/commons/thumb/0/03/Turing_Machine_Model_Davey_2012.jpg/1280px-Turing_Machine_Model_Davey_2012.jpg",
    image_alt = "Physical Turing Machine",
    og_author = c("David J. Hunter")
  )
```

This site provides slides, assignments, and other resources that I use in my class on the theory of computation (CS/MA-135: Formal Languages and Automata, Westmont College).

# Slides 

1. [Introduction to Automata](slides/s1_1finite_automata.html) 
2. [Regular Operations; Nondeterminism](slides/s1_1Breg_ops.html)
3. [Nondeterminism and Closure](slides/s1_2Anodeterminism.html)

# Prework

1. [Finite Automata](prework/p11Aautomata.pdf)
2. [Regular Operations; NFAs](prework/p11Bregops.pdf)
3. [Nondeterminism](prework/p12Anondeterminism.pdf)

# Other Resources

- [Syllabus](syllabus/csma135hunterS24.pdf) (PDF)

All of the pages and slides on this site were authored in RMarkdown. The source code is available on GitHub: https://github.com/djhunter/computation

