---
title: "Formal Languages and Automata"
author: "David J. Hunter"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(metathis)
meta() %>%
  meta_description(
    "Slides, assignments, and code for my class on the theory of computation: CS/MA-135: Formal Languages and Automata."
  ) %>%
  meta_name("github-repo" = "djhunter/computation") %>%
  meta_viewport() %>%
  meta_social(
    title = "Formal Languages and Automata",
    url = "https://djhunter.github.io/computation/index.html",
    image = "https://upload.wikimedia.org/wikipedia/commons/thumb/0/03/Turing_Machine_Model_Davey_2012.jpg/1280px-Turing_Machine_Model_Davey_2012.jpg",
    image_alt = "Physical Turing Machine",
    og_author = c("David J. Hunter")
  )
```

This site provides slides, assignments, and other resources that I use in my class on the theory of computation (CS/MA-135: Formal Languages and Automata, Westmont College).

# Slides 

1. [Introduction to Automata](slides/s1_1finite_automata.html) 
2. [Regular Operations; Nondeterminism](slides/s1_1Breg_ops.html)
3. [Nondeterminism and Closure](slides/s1_2Anodeterminism.html)
4. [NFA Simulation](slides/s1_2BNFAsimulator.html)
5. [Regular Expressions](slides/s1_3Areg_exp.html)
6. [Converting DFA's to Regular Expressions](slides/s1_3Bdfa_to_reg_exp.html)
7. [Nonregular Languages](slides/s1_4Apumping.html)
8. [Pumping Lemma Proofs](slides/s1_4Bplproofs.html)
9. [Regular Languages Review](slides/s1_4Cproblems.html) <br> Exam #1
10. [Grammars and Context-Free Languages](slides/s2_1Agrammars.html)
11. [Pushdown Automata](slides/s2_2Apushdown.html)
12. [Building a PDA for a CFG](slides/s2_2Bcfgtopda.html)
13. [Pumping Lemma for CFL's](slides/s2_3ApumpingCFL.html)
14. [Turing Machines](slides/s3_1Aturingmachines.html)
15. [Turing Machines, continued](slides/s3_1Bturingmachines.html)
16. [Turing Machine Variants](slides/s32AvariantTMs.html)
17. [The Church-Turing Thesis](slides/s33ACTthesis.html) <br>
17b. [A couple more review problems](slides/s33Breview.html) <br> Exam #2
18. [Decidable Languages](slides/s41Adecidable.html)
19. [Decidable Languages involving Grammars](slides/s41Bcfgdeciders.html)
20. [Uncountablility and Recognizability](slides/s42Aunrecognizable.html)
21. [An Undecidable Language](slides/s42Bundecidable.html)
22. [Reduction Arguments](slides/s51Areduction.html)
23. [Computation Histories](slides/s51Bcomphist.html)
24. [The Post Correspondence Problem](slides/s52Apostcorrprob.html)
25. [Mapping Reducibility](slides/s53Amapreduce.html)
26. [More Mapping Reducibility](slides/s53Bmoremapreduce.html)
27. [Review for Test #3](slides/s53Cch45review.html)

---
# 51B comphist friday 3/22
# 52A post correspondence 3/25
# 53A map reduction 3/27
# 53B more map red 4/3
# 53C chapters 4-5 review 4/5
# exam 3 (chapters 4-5) monday, 4/8
---

# Prework

1. [Finite Automata](prework/p11Aautomata.pdf)
2. [Regular Operations; NFAs](prework/p11Bregops.pdf)
3. [Nondeterminism](prework/p12Anondeterminism.pdf)
4. [Simulating an NFA on a DFA](prework/p12Bnfasimulation.pdf)
5. [Regular Expressions](prework/p13Areg_exp.pdf)
6. [Converting DFA's to Regular Expressions](prework/p13Bdfa_to_reg_exp.pdf)
7. [The Pumping Lemma](prework/p14Apumping.pdf)
8. [More Pumping Lemma Proofs](prework/p14Bmorepump.pdf)
9. [Regular vs. Nonregular](prework/p14Cregprobs.pdf)
10. [Context-Free Grammars](prework/p21Acfg.pdf)
11. [Pushdown Automata](prework/p22Apushdown.pdf)
12. [PDA's from CFG's](prework/p22Bcfgtopda.pdf)
13. [Pumping Lemma for CFL's](prework/p23ApumpingCFL.pdf)
14. [Turing Machines](prework/p31Aturingmachines.pdf)
15. [More Turing Machines](prework/p31Bturingmachines.pdf)
16. [Turing Machine Variants](prework/p32AtmVariants.pdf)
17. [Selected Review Problems](prework/p33ACTthesis.pdf)
18. [Decidable Languages](prework/p41Adecideable.pdf)
19. [Countable Sets](prework/p41Bcountable.pdf)
20. [Uncountability and non-recognizable languages](prework/p42Aunrecognizable.pdf)
21. [More decidability and recognizability problems](prework/p42Bdecidability.pdf)
22. [Reduction Arguments](prework/p51Areductions.pdf)
23. [Arguments using Computation Histories](prework/p51Bcomphist.pdf)
24. [The Post Correspondence Problem](prework/p52Apostcorprob.pdf)
25. [Mapping Reducibility](prework/p53Amapreduce.pdf)
26. [Mapping reducibility and recognizability proofs](prework/p53Bmoremapreduce.pdf)

---
---

# Other Resources

- [Syllabus](syllabus/csma135hunterS24.pdf) (PDF)

All of the pages and slides on this site were authored in RMarkdown. The source code is available on GitHub: https://github.com/djhunter/computation

