---
title: "Section 1.3: Regular Expressions"
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: notesheader.tex
  revealjs::revealjs_presentation:
    center: no
    css: slidesdjh.css
    fig_height: 7
    fig_width: 14
    highlight: zenburn
    reveal_options:
      controls: yes
      height: 600
      progress: no
      width: 1080
    reveal_plugins: notes
    self_contained: no
    theme: black
    transition: slide
  html_document:
    self_contained: true
    theme: spacelab
    code_download: true
date: "January 15, 2020"
---

```{r include=FALSE}
set.seed(11320)
fnames <- sample(c("Sam", "James", "Drake", "Wesley", "Sophia", "Carolyn", "Trevor", "Tristan", "Ryley"))
```

# Prework Presentations

## 1. [`r paste(fnames[1:3], sep = ',')`] 

Apply the construction of an NFA simulator (Theorem 1.39) to obtain a DFA that is equivalent to the following NFA.

```{r, echo=FALSE, fig.width=5, fig.height=4, message=FALSE, fig.align='center', warning=FALSE}
library(igraph)
a3 <- graph_from_literal(" "-+q1-+q2-+q1, q3-+q2, q1-+q3-+q3, simplify=FALSE)
#coords <- matrix(c(-3,0, 0,0, 3,0, 1.5, -1), nrow = 4, byrow=TRUE)
coords <- layout_as_tree(a3)
plot(a3, 
     edge.curved = c(0, 0.5, 0.5, 0, 0, 0.5), 
     layout=coords,
     vertex.size = 25,
     vertex.shape = c("none", "circle", "circle", "circle"),
     vertex.label = c(" ", expression(italic(q)[1]), expression(italic(q)[2]), expression(italic(q)[3])),
     edge.loop.angle=3.14*0,
     mark.groups = list(3), mark.shape = .6,  mark.col = "pink", mark.border = "black",
     vertex.label.cex = 1,
     edge.label = c("", expression(atop(epsilon,  " ")),"a         ","a, b         \n","      a","    b"), 
     edge.label.cex = 1,
     edge.width = 2,
     edge.arrow.size=0.7,
     ylim = c(-1.02,0.2), xlim=c(-1,1.3))
```


## 2. [`r paste(fnames[4:6], sep = ',')`] 

Determine the language recognized by the following NFA, and express your answer using the $*$ notation.

```{r, echo=FALSE, fig.width=5, fig.height=5, message=FALSE, fig.align='center', warning=FALSE}
library(igraph)
a3 <- graph_from_literal(" "-+q1-+q2-+q1, q2-+q3-+q1, q2-+q4-+q1, simplify=FALSE)
#coords <- matrix(c(-3,0, 0,0, 3,0, 1.5, -1), nrow = 4, byrow=TRUE)
coords <- layout_as_tree(a3)
plot(a3, 
     edge.curved = c(0, 0.5, 0.5, 0, 0, 0, 0), 
     layout=coords,
     vertex.size = 35,
     vertex.shape = c("none", "circle", "circle", "circle", "circle"),
     vertex.label = c(" ", expression(italic(q)[1]), expression(italic(q)[2]), expression(italic(q)[3]), expression(italic(q)[4])),
     mark.groups = list(2), mark.shape = -0.9,  mark.col = "pink", mark.border = "black",
     vertex.label.cex = 1, 
     edge.label = c("", "   0", "\n1  ", "\n0", "1\n", "\n1", "     0"), 
     edge.label.cex = 1,
     edge.width = 2,
     edge.arrow.size = 0.5)
#     ylim = c(-1.02,0.2), xlim=c(-1,1.3))
```

## 3. [`r paste(fnames[7:9], sep = ',')`] 

Apply the construction of an NFA simulator (Theorem 1.39) to obtain a DFA that is equivalent to the NFA given in Problem 2. Simplify if possible.

```{r, echo=FALSE, fig.width=5, fig.height=5, message=FALSE, fig.align='center', warning=FALSE}
library(igraph)
a3 <- graph_from_literal(" "-+q1-+q2-+q1, q2-+q3-+q1, q2-+q4-+q1, simplify=FALSE)
#coords <- matrix(c(-3,0, 0,0, 3,0, 1.5, -1), nrow = 4, byrow=TRUE)
coords <- layout_as_tree(a3)
plot(a3, 
     edge.curved = c(0, 0.5, 0.5, 0, 0, 0, 0), 
     layout=coords,
     vertex.size = 35,
     vertex.shape = c("none", "circle", "circle", "circle", "circle"),
     vertex.label = c(" ", expression(italic(q)[1]), expression(italic(q)[2]), expression(italic(q)[3]), expression(italic(q)[4])),
     mark.groups = list(2), mark.shape = -0.9,  mark.col = "pink", mark.border = "black",
     vertex.label.cex = 1, 
     edge.label = c("", "   0", "\n1  ", "\n0", "1\n", "\n1", "     0"), 
     edge.label.cex = 1,
     edge.width = 2,
     edge.arrow.size = 0.5)
#     ylim = c(-1.02,0.2), xlim=c(-1,1.3))
```

# Languages and Automata

## So far...

- **Regular Languages** are languages that are recognized by DFA's
- We have proved:
    + DFA's can simulate NFA's, so we can use NFA's instead.
    + Regular languages are closed under **regular operations** $\cup$, $\circ$, $*$

# Regular Expressions

## Describing Languages

So far, we have specified particular languages a couple ways:

- Set builder notation: $\{w \mid \text{some property of } w\}$
- The language recognized by a given machine.

Today we will describe them using **regular expressions.**

## Example: Star notation

Star notation was handy for expressing "give me all the ways to combine this set of strings."

- Prework #2
- $\{0 ,1\}^*$

Regular expressions allow concatination, union, and star.

## Definition of a Regular Expression

$R$ is a **regular expression** if $R$ is:

1. $a$, for some $a\in \Sigma$,
2. $\varepsilon$
3. $\emptyset$
4.  $R_1 \cup R_2$, where $R_1$ and $R_2$ are regular expressions.
5. $R_1\circ R_2$,  where $R_1$ and $R_2$ are regular expressions.
6. $R_1^*$, where $R_1$ is a regular expression.

Notice that this is a **recursive definition**.

## Interpreting Regular Expressions

Let $\Sigma = 0\cup 1 = \{0,1\}$. What languages are these?

1. $1\Sigma$ *$\quad$ (This is how we write $1\circ \Sigma$.)*
2. $1\Sigma^*$
3. $(\Sigma\Sigma)^*$
4. $1^*01^+$ *$\quad$ ($a^+$ is shorthand for $aa^*$)*
5. $1^*(01^+)^*$
6. $(0\cup\varepsilon)1$
7. $\Sigma\emptyset$
8. $\emptyset^*$

## Writing Regular Expressions

Which regular expressions define these languages over $\Sigma = \{0,1\}$?

1. $\{w \mid w \text{ starts with a 1 and ends with a 0} \}$
2. $\{w \mid w \text{ contains 01 as a substring} \}$
3. $\{w \mid w \text{ contains an odd number of 1's} \}$

# Are Regular Expressions the same as regular languages?

## Regular Expression $\Rightarrow$ Regular Language

Lemma 1.55. If a language is described by a regular expression, then it is a regular language.

Proof: Use induction on the recursive definition.

## Use the proof to construct an NFA

Follow the logic of the proof to build an NFA that recognizes $(ab \cup a)^*$.

# Applications

## Some examples of using grep

https://www.digitalocean.com/community/tutorials/using-grep-regular-expressions-to-search-for-text-patterns-in-linux

## Lexical Analyzers ("lexers")

Numerical constants can be integers, decimals, possibly with leading 0's, possibly with +/- signs.

- $D = \{0,1,2,3,4,5,6,7,8,9\}$
- $\Sigma = D \cup \{+, -, .\}$

Numerical constant:

$$
(+ \cup - \cup \varepsilon)(D^+ \cup D^+.D^* \cup D^*.D^+)
$$
