---
geometry: left=0.75in, right=0.75in, top=0.8in, bottom=0.4in
classoption: twoside
output: 
  pdf_document:
    keep_tex: true
    includes:
      in_header: preworkheader.tex
---

```{r include=FALSE}
set.seed(1620)
fnames <- sample(c("Sam", "James", "Drake", "Wesley", "Sophia", "Carolyn", "Trevor", "Tristan", "Ryley"))
```

## Prework 1.1a: Finite Automata

Write your preliminary solutions to each problem on this sheet of paper (front and back). In addition, you may record any important facts and definitions, and include any questions that you have. Add additional sheets as necessary. For written prework, everyone should do every problem. The names in brackets indicate who is responsible for presenting the problem; one of these students will be chosen to explain the problem in class.

1. [`r paste(fnames[1:3], sep = ',')`] For the finite automata with the following state diagram, identify $Q$, $\Sigma$, $\delta$, $q_0$, and $F$, and describe the language it recognizes.

```{r, echo=FALSE, fig.width=8, fig.height=3.5, message=FALSE, fig.align='center'}
library(igraph)
a3 <- graph_from_literal(" "-+q1-+q2-+q3-+q2-+q2, q1-+q1, q2-+q2, simplify=FALSE)
coords <- layout_on_grid(a3, width = 4)
plot(a3, 
     edge.curved = c(0, 0, 0.5, 0.5, 0.5, 0.5), 
     layout=coords,
     vertex.size = 15,
     vertex.shape = c("none", "circle", "circle", "circle"),
     vertex.label = c(" ", expression(italic(q)[1]), expression(italic(q)[2]), expression(italic(q)[3])),
     edge.loop.angle=3.14*1.5,
     mark.groups = list(3), mark.shape = 0.75, mark.col = "pink", mark.border = "black",
     vertex.label.cex = 1,
     edge.label = c("","     1\n","0\n","  0,1\n","1\n","0\n"), 
     edge.label.cex = 1,
     edge.width = 2,
     ylim = c(-1.1,-0.6), xlim=c(-0.5,1))
```


2. [`r paste(fnames[4:6], sep = ',')`] Here is question 2. It has an illustration.

3. [`r paste(fnames[7:9], sep = ',')`] This is question 3.

\medskip
\mbox{}\xrfill[2pt]{1pt}\textsc{\small begin your solutions below this line}\xrfill[2pt]{1pt}\mbox{}

