---
geometry: left=0.75in, right=0.75in, top=0.8in, bottom=0.4in
classoption: twoside
output: 
  pdf_document:
    keep_tex: false
    includes:
      in_header: preworkheader.tex
---

```{r include=FALSE}
set.seed(11624)
fnames <- sample(c("Ky", "Curtis", "Micah", "Todd", "Andrew", "Meghan", "Allie", "David", "Connor", "Levi", "Ben", "Grace", "Joshua"))
```

## Prework 1.2b: Simulating an NFA on a DFA

Write your preliminary solutions to each problem on this sheet of paper (front and back). The names in brackets indicate the subset responsible for presenting the problem.

1. [`r paste(fnames[1:4], sep = ',')`] Apply the construction of an NFA simulator (Theorem 1.39) to obtain a DFA that is equivalent to the following NFA.

```{r, echo=FALSE, fig.width=5, fig.height=4, message=FALSE, fig.align='center', warning=FALSE}
library(igraph)
a3 <- graph_from_literal(" "-+q1-+q2-+q1, q3-+q2, q1-+q3-+q3, simplify=FALSE)
#coords <- matrix(c(-3,0, 0,0, 3,0, 1.5, -1), nrow = 4, byrow=TRUE)
coords <- layout_as_tree(a3)
plot(a3, 
     edge.curved = c(0, 0.5, 0.5, 0, 0, 0.5), 
     layout=coords,
     vertex.size = 25,
     vertex.shape = c("none", "circle", "circle", "circle"),
     vertex.label = c(" ", expression(italic(q)[1]), expression(italic(q)[2]), expression(italic(q)[3])),
     edge.loop.angle=3.14*0,
     mark.groups = list(3), mark.shape = .6,  mark.col = "pink", mark.border = "black",
     vertex.label.cex = 1,
     edge.label = c("", expression(atop(epsilon,  " ")),"a         ","a, b         \n","      a","    b"), 
     edge.label.cex = 1,
     edge.width = 2,
     edge.arrow.size=0.7,
     ylim = c(-1.02,0.2), xlim=c(-1,1.3))
```


2. [`r paste(fnames[5:8], sep = ',')`] Determine the language recognized by the following NFA, and express your answer using the $*$ notation.

```{r, echo=FALSE, fig.width=5, fig.height=4, message=FALSE, fig.align='center', warning=FALSE}
library(igraph)
a3 <- graph_from_literal(" "-+q1-+q2-+q1, q2-+q3-+q1, q2-+q4-+q1, simplify=FALSE)
#coords <- matrix(c(-3,0, 0,0, 3,0, 1.5, -1), nrow = 4, byrow=TRUE)
coords <- layout_as_tree(a3)
plot(a3, 
     edge.curved = c(0, 0.5, 0.5, 0, 0, 0, 0), 
     layout=coords,
     vertex.size = 35,
     vertex.shape = c("none", "circle", "circle", "circle", "circle"),
     vertex.label = c(" ", expression(italic(q)[1]), expression(italic(q)[2]), expression(italic(q)[3]), expression(italic(q)[4])),
     mark.groups = list(2), mark.shape = -0.9,  mark.col = "pink", mark.border = "black",
     vertex.label.cex = 1, 
     edge.label = c("", "   0", "\n1  ", "\n0", "1\n", "\n1", "     0"), 
     edge.label.cex = 1,
     edge.width = 2,
     edge.arrow.size = 0.5)
#     ylim = c(-1.02,0.2), xlim=c(-1,1.3))
```

3. [`r paste(fnames[9:13], sep = ',')`] Apply the construction of an NFA simulator (Theorem 1.39) to obtain a DFA that is equivalent to the NFA given in Problem 2. Simplify if possible.

\medskip
\mbox{}\xrfill[2pt]{1pt}\textsc{\small begin your solutions below this line}\xrfill[2pt]{1pt}\mbox{}

