---
geometry: left=0.75in, right=0.75in, top=0.8in, bottom=0.4in
classoption: twoside
output: 
  pdf_document:
    keep_tex: false
    includes:
      in_header: preworkheader.tex
---

```{r include=FALSE}
set.seed(11924)
fnames <- sample(c("Ky", "Curtis", "Micah", "Todd", "Andrew", "Meghan", "Allie", "David", "Connor", "Levi", "Ben", "Grace", "Joshua"))
```

## Prework 1.3b: Converting DFA's to Regular Expressions

Write your preliminary solutions to each problem and submit a PDF on Canvas. The names in brackets indicate the subset responsible for presenting the problem.

1. [`r paste(fnames[1:3], sep = ',')`] Use the method described in the proof of Lemma 1.60 to convert the following DFA into a regular expression. (First convert the DFA to a GNFA, then delete two states to obtain a GNFA with a single edge giving the regular expression.)

```{r, echo=FALSE, fig.width=5, fig.height=6, message=FALSE, fig.align='center'}
library(igraph)
a3 <- graph_from_literal(" "-+q1-+q2-+q1-+q1, q2-+q2, simplify=FALSE)
coords <- layout_on_grid(a3, width = 3)
plot(a3, 
     edge.curved = c(0, 0.5, 0.5, 0.5, 0.5), 
     layout=coords,
     vertex.size = 25,
     vertex.shape = c("none", "circle", "circle"),
     vertex.label = c(" ", expression(italic(q)[1]), expression(italic(q)[2])),
     edge.loop.angle=3.14*1.5,
     mark.groups = list(3), mark.shape = 0.7, mark.col = "pink", mark.border = "black",
     vertex.label.cex = 1,
     edge.label = c("","    b\n","  b\n"," a\n"," a\n"), 
     edge.label.cex = 1,
     edge.width = 2)
```

2. [`r paste(fnames[4:6], sep = ',')`] Use the method described in the proof of Lemma 1.60 to convert the following DFA into a regular expression. 

```{r, echo=FALSE, fig.width=5, fig.height=7, message=FALSE, fig.align='center', warning=FALSE}
library(igraph)
a3 <- graph_from_literal(" "-+q1-+q2-+q3-+q1, q3-+q2-+q2, simplify=FALSE)
#coords <- matrix(c(-3,0, 0,0, 3,0, 1.5, -1), nrow = 4, byrow=TRUE)
set.seed(3321)
coords <- layout_nicely(a3)
plot(a3, 
     edge.curved = c(0, 0, 0.5, 0, 0.5, 0.5), 
     layout=coords,
     vertex.size = 25,
     vertex.shape = c("none", "circle", "circle", "circle"),
     vertex.label = c(" ", expression(italic(q)[1]), expression(italic(q)[2]), expression(italic(q)[3])),
     edge.loop.angle=3.14*0,
     mark.groups = list(2,4), mark.shape = .6,  mark.col = "pink", mark.border = "black",
     vertex.label.cex = 1,
     edge.label = c("", "        a,b", "b\n", "a      \n", "\nb", "    a"), 
     edge.label.cex = 1,
     edge.width = 2,
     edge.arrow.size=0.7,
     ylim = c(-1.02,0.2), xlim=c(-1,1.3))
```

\medskip 
\textsl{For problems 3 and 4 we are using the notation $a^k$ to represent the string $\underbrace{aaa \cdots a}_k$.}

3. [`r paste(fnames[7:9], sep = ',')`] Let $A = \{0^n1^n \mid  n \in \mathbb{N}\}$. Which of the following regular expressions describe the language $A$? For each expression, explain why the expression describes $A$, or why it doesn't.
$$
0^*1^*, \quad 0^+1^+, \quad 0^*011^*, \quad (01)^*, \quad (01)^+ 
$$

4. [`r paste(fnames[10:13], sep = ',')`] Suppose that $M$ is a DFA with 5 states that accepts the string \texttt{000000}.
    a. Explain why, when $M$ processes \texttt{000000}, some state must occur twice.
    b. Explain why $M$ must also accept infinitely many strings of the form
    $\mathtt{0}^k$ for certain values of $k > 6$.
    
\medskip
\mbox{}\xrfill[2pt]{1pt}\textsc{\small begin your solutions below this line}\xrfill[2pt]{1pt}\mbox{}

    